---
title: "special_topics_1"
author: "Caleigh Dwyer"
date: "2023-11-28"
output: html_document
---

homework6: out of 1000 bootstraps how many were NAs because they were negative and then graph those that are nonnegative

created r package to help hide data folder from github directory when working with patient data. adds a little arrow to the data directory, making it not a folder in the project, it is just a symbolic link. git doesnt look inside of that data folder. as long as working on your computer, you can access and work with it, but others cannot. R recognizes aliases but git ignores them

you can cache results, which is useful if you're running big analyses cuz it saves the results rather than requiring that it be re run each time the output is made.

sometimes you need address data anyways to run a neighborhood analysis/geo analysis, but this is still identifiable information so you'll need to store it carefully

DStore you probs dont need to commit them to your github because it's just about Mac's system architecture. 

```{r}
library(tidyverse)
```

```{r}
litters_df = 
  read_csv(here::here("data", "FAS_litters.csv"))
```

## Shiny

Framework for building interactive plots and web apps in r

Allows you to create a graphical user interface (GUI). Users can interact w/ your code w/o knowing r, and you can communicate viz, models, algorithms w/ collaborators

uses html, css, and javascript framework (you don't need to know these to use shiny tho). This was developed in 2012.

adds interacctivity -- your app can take user input and update outputs accordingly

run this in console to see an example: > shiny::runExample("01_hello")

sharing shiny apps is more complex than just creating a dashboard. you have to deploy in special ways, not just github cuz you need r to run in the background

# inputs
Widgets are text elements that users can interact w/  (scroll bars, buttons, text, slider, radio buttons)

functions that react to user input:
renderPrint
renderText
renderTable
renderPlot
renderPlotly

there are standalone shiny apps, not built within an R markdown document. Can be more flexible than piggybacking on R markdown/flexdash

shiny apps:
-ui: controls layout and appearance, where you add widgets
-server

HTML files can't do this. You have to send raw files to a host and then recipient knits the file/runs the app through rstudio
(shinyapps.io is pretty common)


# Working on an example
use "run document" for the template rather than knit this

radioButtons(
  inputID = "room_type",
  label = h3("Select a price range"),
  choices = room_vec,
  select = "Entire home/apt"
)

renderText({input[["price_range"]]})

